---
title: "Assignment 1"
author: "Joseph Kambourakis"
date: "Friday, March 06, 2015"
output: html_document
---

#Loading and Preprocessing the Data

```{r, echo = T}
activity = read.csv('activity.csv')
library(plyr)
```

## 1.  The Total Number of Steps Taken 

```{r}
a =ddply(activity, ~date, summarise, Total =sum(steps, na.rm = T))
print(a)
```

2.  Histogram of the Total Number of Steps Taken Each Day
```{r, echo = F}
hist(a$Total, xlab = "Total Number of Steps", main = "Histogram of Total Number of Steps taken each day")
```

3.  Mean Number of Steps
```{r}
mean(a$Total, na.rm = T)
```
4.  Median Number of Steps
```{r}
median(a$Total, na.rm=T)
```

#What is the average daily activity pattern?

Time series plot of the average number of steps taken (averaged across all days) versus the 5-minute intervals?
```{r}
b =ddply(activity, ~interval, summarise, Mean =mean(steps, na.rm = T))
plot(b, type = "l", ylab = "Average Across All Days")
```

Which 5 minute interval contains the most steps?
```{r}
b$interval[b$Mean==max(b$Mean, na.rm = T)]
```

Imputing Missing Values  
1. Calculate the Total number of missing values
```{r}
count(activity$steps == "NA")[2,2]
```

2.  Stragegy for filling in data set
```{r}
activity$steps[is.na(activity$steps)]<-mean(activity$steps, na.rm = T)
```

3.  Create a new dataset named cleanedactivity
```{r}
activity$steps[is.na(activity$steps)]<-mean(activity$steps, na.rm = T)
cleanedactivity = activity
```

4.  Histogram of the total number of steps taken each day after missing values were imputed
```{r}
c =ddply(cleanedactivity, ~date, summarise, Total =sum(steps, na.rm = T))
hist(a$Total, xlab = "Total Number of Steps", main = "Histogram of Cleaned Number of Steps taken each day")
mean(c$Total, na.rm = T)
median(c$Total, na.rm = T)
```


Are there Differences between Weekdays and Weekends

1. Create a new factor variable in the dataset with two levels
```{r}
cleanedactivity$Day = (weekdays(as.Date(cleanedactivity$date)))
cleanedactivity$Day[cleanedactivity$Day == "Sunday" | cleanedactivity$Day == "Saturday" ]="weekend"
cleanedactivity$Day[cleanedactivity$Day == "Monday" | cleanedactivity$Day == "Tuesday"| cleanedactivity$Day == "Wednesday"| cleanedactivity$Day == "Thursday"| cleanedactivity$Day == "Friday"]="weekday"
cleanedactivity$Day = factor(cleanedactivity$Day)
```

2. Make a panel plot containing a time series plot
```{r}
d =ddply(cleanedactivity, .(interval,Day), summarise, Mean =mean(steps, na.rm = T))
we=subset(d, Day == "weekend")
wd = subset(d, Day =="weekday")
library(lattice)
xyplot(Mean~interval|Day, 
           data = d,
           type = "l",
           xlab = "Interval",
           ylab = "Number of steps",
           layout=c(1,2))
```